### 2020-10-01 Paper/active-learning-interactive.tex
added active-learning-interactive tex file

Recent work in machine translation approaches has focused on integrating active learning
with interaction-based neural machine translation. These modifications have been experimentally 
employed with unbounded real-time streaming data \cite{peris2020active} with the objective of selecting 
only those sentences which are very difficult for the translation system to accurately classify upon 
during the initial stages. Results have shown promise in improving the overall quality of translation 
and reducing human intervention in the process. A core component of this approach involves tweaking the 
sampling strategies involved to maximize the probability of selecting only the most “difficult” sentences. 
However, these results have a lack of representation and probable efficacy with respect to low-resource-based 
multilingual paradigms. The effectiveness of slight alterations in the overall encoder-decoder transformer 
network also needs to be experimented upon to possibly generate higher quality translations.