### 2020-10-10 Paper/active-learning.tex
updated related work

Active Learning has been widely studied in Natural Language Processing and often integrated as an effective approach in low resource scenaria in modern neural network systems and for a variety of tasks. 
\citet{lu2020investigating} suggest that representations generated from pre-trained transformer neural network models, are more effective in Active Learning for text classification, than using simpler word embeddings.

Closer to our Machine Translation objective, \citet{zhang2018active} propose two methods, for sentence selection based on considering semantic similarity between sentence embeddings or decoder probabilities as tokens of translation quality, while \citet{peris2020active} explore the application of active learning techniques in the interactive translation of large data streams. The latter work uses the attention mechanism of a neural machine translation system to query only for those sentences worthy of human annotation for translation.

In a more recent approach, \citet{zeng2019empirical} perform an empirical evaluation of different AL methods for Transformers \cite{vaswani2017attention}, analyzing data-driven approaches, that sample instances using two sets of labeled and unlabeled sentences, and model-driven methods, which additionally incorporate information from the current model into sampling. However their work notes the weak performance of Coverage Sampling compared to the other techniques, due to the complex nature of the state-of-the art multi-layer multi-head Natural Language Processing architectures, a phenomenon we are addressing with the current work.

A line of work has also explored the shortcomings of Active Learning approaches in Machine Translation. \citet{lowell2018practical} observe the performance drop issues arising across different models and tasks and discuss the methods' generalization constraints. Similarly, \citet{koshorek2019limits} examine the potential and limitations of an oracle policy in selecting samples to maximize the performance of a neural machine translation system in learning sentence semantic representations.

%Recent work in machine translation approaches has focused on integrating active learning
%with interaction-based neural machine translation. These modifications have been experimentally 
%employed with unbounded real-time streaming data \cite{peris2020active} with the objective of selecting 
%only those sentences which are very difficult for the translation system to accurately classify upon %
%during the initial stages. Results have shown promise in improving the overall quality of translation 
%and reducing human intervention in the process. A core component of this approach involves tweaking the 
%sampling strategies involved to maximize the probability of selecting only the most “difficult” sentences. 
%However, these results have a lack of representation and probable efficacy with respect to low-resource-based 
%multilingual paradigms. The effectiveness of slight alterations in the overall encoder-decoder transformer 
%network also needs to be experimented upon to possibly generate higher quality translations. s