### 2020-11-12 Paper/datasets.tex
added fce dataset info

To train our classifier on the target language (English), we explored three
different datasets, namely: (1) Cambridge FCE Corpus; (2) CoNLL-14; (3) CoNLL-15

\paragraph{Cambridge FCE Corpus.}
The Cambridge First Certificate Exam Learner Corpus is a 16-million word 
corpus of English examination scripts collected by Cambridge University Press. 
The scripts categorize errors in the sentences into approximately 80 error labels, 
along with corrections \cite{NichollsFce}.

\paragraph{CoNLL-13.}

\paragraph{CoNLL-14.}