### 2020-11-13 Paper/data-processing.tex
more updates on the paper for error classification

We divided data processing into \emph{x} steps -
    (1) Mapping different error labels to five main error labels;
    (2) Combining the dataset with both correct and incorrect sentences; 
    (3) Encoding error labels and building an error vector;
    (4) Ensuring sentence "uniqueness"; and

\subparagraph{Mapping Error Labels.}
The approximately 80 different error labels were mapped to the their five higher 
levels manually. \cite{NichollsFce} describes the tagging process for the error labels, 
where tags starting with "F" indicated incorrect form, "M" indicated missing, "R" indicated replacing,
"U" undicated unnecessary, etc. Further, the taxonomy of translation errors from \cite{costaTaxonomyErrorLabels}
was used for the appropriate mappings.

\subparagraph{Combining dataset with correct and incorrect sentences.}
The FCE text corpus provides the examination scripts in an XML format. These are initially converted to "txt" files, 
and then parsed into a comma separated format (csv) with every sentence/group of sentences with their error tags. 
In case of multiple errors per sample, the sentence/group is repeated for every tag in the csv file. 
Since the Cambridge FCE corpus provided us with error tags for sentences along with their 
corrections, the correct sentences are filtered out separately. Further, an additional "CO" label denoting correctness
is attached to each of these sentences and then are appended to the dataset containing incorrect ones.

\subparagraph{Encoding Error Labels \& Error vector.}
Classifying sentences grouped into six categories (5 error labels and 1 correct label) looks to be a multi-class 
classification problem, however, it is possible that a given sentence or group of sentences might have multiple errors 
in them. Thus, every sample's error tag is one-hot-encoded: six new columns are added, each for a given error tag, with 
a "1" on the actual error tag's column. An error vector is built using these one-hot-encodings, each vector for a given sample 
of length equalling 6. As sentences can have multiple errors, those samples with repeating rows are grouped, and their 
error vectors are aggregated.

\subparagraph{Sentence Uniqueness.}
Further, it is possible to have a sentence that is correct in one context, but has errors in 
another context (a simple example would be closing statements). For this reason, those samples that have a "1" for both "CO" and 
other error labels are dropped from the dataset.